On the surprising effectiveness of a simple matrix exponential derivative approximation, with application to global SARS-CoV-2

### 2023-10-25 manuscript/pnas_revision.tex
edits to caption

\begin{figure*}[!t]
	\centering
	\includegraphics[width=\linewidth]{truthMeanFig.pdf}
	\caption{Posterior means versus truth for CTMC generator matrix elements under differing sparsity regimes and different dimensionalities $d$ holding the number of observations fixed at 300.  To affect sparsity levels, we generate generator matrix entries according to the Bayesian bridge distribution \cite{polson2014bayesian} with different exponents ($\alpha \in \{1,1/2,1/4\}$), normalizing by the largest absolute values to ease comparison. Smaller $\alpha$ values encode more peaked distributions with heavier tails and thus enforce greater sparsity.   Plot \textbf{c} reflects the fact that the Bayesian bridge prior with exponent $\alpha=1/4$ helps identify non-null parameters in small sample contexts \cite{magee2023random}.  With this intuition in mind, we specify such a prior on generator random effects in \cref{sec:sars:cov:app}.}
	\label{fig:truthMean}
\end{figure*}

>>>>

\begin{figure*}[!t]
	\centering
	\includegraphics[width=\linewidth]{truthMeanFig.pdf}
	\caption{Posterior means versus truth for CTMC generator matrix elements within differing sparsity regimes and with different dimensionalities $d$, holding observation count fixed at 300. We generate posteriors using surrogate-trajectory Hamiltonian Monte Carlo with the naive matrix exponential derivative.  To affect sparsity levels, we generate generator matrix entries according to the Bayesian bridge distribution \cite{polson2014bayesian} with different exponents ($\alpha \in \{1,1/2,1/4\}$), normalizing by the largest absolute values to ease comparison. Smaller $\alpha$ values encode more peaked distributions with heavier tails and thus enforce greater sparsity.   Plot \textbf{c} reflects the fact that the Bayesian bridge prior with exponent $\alpha=1/4$ helps identify non-null parameters in small sample contexts \cite{magee2023random}.  With this intuition in mind, we specify such a prior on generator random effects in \cref{sec:sars:cov:app}.}
	\label{fig:truthMean}
\end{figure*}
